---
title: "Hands-on Exercise 4"
execute:
  eval: true
  echo: true
  warning: false
---

# Getting Started

## Installing Packages

For this exercise, the following R packages will be used:

-   tidyverse, a family of R packages for data science process.

-   ggridges, a ggplot2 extension specially designed for plotting ridgeline plots.

-   ggdist for visualising distribution and uncertainty.

-   ggstatsplot package to create visual graphics with rich statistical information.

-   performance package to visualise model diagnostics.

-   parameters package to visualise model parameters.

-   plotly for creating interactive plot.

-   gganimate for creating animation plot.

-   DT for displaying interactive html table.

-   crosstalk for for implementing cross-widget interactions (currently, linked brushing and filtering).

-   readr for importing csv into R.

-   FunnelPlotR for creating funnel plot.

-   ggplot2 for creating funnel plot manually.

-   knitr for building static html table.

```{r}
devtools::install_github("wilkelab/ungeviz")
pacman::p_load(ggdist, ggridges, ggthemes, colorspace, tidyverse, ggstatsplot, rstantools, readxl, performance, parameters, see, ungeviz, plotly, crosstalk, DT, gganimate, FunnelPlotR, knitr)
```

## Data Import

Once again, Exam_data.csv will be utilized for the exercise.

```{r}
exam <- read_csv("data/Exam_data.csv")
```

Once these steps are completed, the exercise may begin.

# Data Visualization

## Ridgeline Plot 

Ridgeline plots show a dataset's distribution of numeric values. There are many options as to how these can be visualized.

::: panel-tabset
### Color Gradient

```{r}
ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS,
           fill = stat(x))) +
  geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Temp. [F]",
                       option = "C") +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

### Probability per Category (Section)

```{r}
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = 0.5 - abs(0.5-stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability",
                       direction = -1) +
  theme_ridges()
```

### Quantile Ranges

```{r}
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = 4,
    quantile_lines = TRUE) +
  scale_fill_viridis_d(name = "Quartiles") +
  theme_ridges()
```

### Quantile with Cut Points

```{r}
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = c(0.025, 0.975)
    ) +
  scale_fill_manual(
    name = "Probability",
    values = c("#FF0000A0", "#A0A0A0A0", "#0000FFA0"),
    labels = c("(0, 0.025]", "(0.025, 0.975]", "(0.975, 1]")
  ) +
  theme_ridges()
```
:::

## Raincloud Plot

Raincloud is a similar visualization technique to that of the boxplot. However, this method allows viewers to gain a clearer view of the dataset by showing areas where densities are clustered. However, a rancloud may be combined with a boxplot and a half-dotplot in order to show a clearer view of the data.

::: panel-tabset
### Lone Raincloud

```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  coord_flip() +
  theme_economist()
```

### Raincloud + Boxplot + Half-dot

```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 1.5) +
  coord_flip() +
  theme_economist()
```
:::

# Visual Statistical Analysis

## One-sample Test

The code chunk below is used to build a one-sample test on the different test scores

::: panel-tabset
### English

```{r}
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

### Math

```{r}
set.seed(1234)

gghistostats(
  data = exam,
  x = MATHS,
  type = "bayes",
  test.value = 60,
  xlab = "Math scores"
)
```

### Science

```{r}
set.seed(1234)

gghistostats(
  data = exam,
  x = SCIENCE,
  type = "bayes",
  test.value = 60,
  xlab = "Science scores"
)
```
:::

## Two-sample Mean Test

The code chunk below builds a visual two-sample mean test for each subject score by gender.

::: panel-tabset
### English

```{r}
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = ENGLISH,
  type = "np",
  messages = FALSE
)
```

### Math

```{r}
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

### Sicence

```{r}
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = SCIENCE,
  type = "np",
  messages = FALSE
)
```
:::

## One-way ANOVA Test

The code chunk below builds a One-way ANOVA test for the subject scores by race.

::: panel-tabset
### English

```{r}
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

### Math

```{r}
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = MATHS,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

### Science

```{r}
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = SCIENCE,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```
:::

## Significant Test of Correlation

The code chunk below can be used to create a Significant Test of Correlation between the scores per subject.

::: panel-tabset
### English-Math

```{r}
ggscatterstats(
  data = exam,
  x = ENGLISH,
  y = MATHS,
  marginal = FALSE,
  )
```

### English-Science

```{r}
ggscatterstats(
  data = exam,
  x = ENGLISH,
  y = SCIENCE,
  marginal = FALSE,
  )
```

### Math-Science

```{r}
ggscatterstats(
  data = exam,
  x = MATHS,
  y = SCIENCE,
  marginal = FALSE,
  )
```
:::

## Significant Test of Association

The code chunk below bins the Math scores into 4-class variables and then builds a visual for Significant Test of Association

```{r}
exam1 <- exam %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)
ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```

# Visualizing Models

For this section, the Toyota Corolla case will be used.The code chunk below imports the data into R.

```{r}
car_resale <- read_xls("data/ToyotaCorolla.xls", 
                       "data")
car_resale
```

The following code chunks are used to check various tests on the data.

::: panel-tabset
### Multicolinearity

```{r}
model <- lm(Price ~ Age_08_04 + Mfg_Year + KM + 
              Weight + Guarantee_Period, data = car_resale)
check_c <- check_collinearity(model)
plot(check_c)
```

### Normality Assumption

```{r}
model1 <- lm(Price ~ Age_08_04 + KM + 
              Weight + Guarantee_Period, data = car_resale)
check_n <- check_normality(model1)
plot(check_n)
```

### Homogeneity of Variances

```{r}
model1 <- lm(Price ~ Age_08_04 + KM + 
              Weight + Guarantee_Period, data = car_resale)
check_h <- check_heteroscedasticity(model1)
plot(check_h)
```
:::

All of these can also be done simultaneously using the code chunk below.

```{r}
check_model(model1)
```

## Visualizing Regression Parameters

There are 2 options to visualize the parameters of a regression model.

::: panel-tabset
### Option 1

```{r}
plot(parameters(model1))
```

### Option 2

```{r}
ggcoefstats(model1, 
            output = "plot")
```
:::

# Visualizing Uncertainty

For this section, we will once again use the Exam Data.

The summary statistics of the data can be obtained using the code below.

```{r}
my_sum <- exam %>%
  group_by(RACE) %>%
  summarise(
    n=n(),
    mean=mean(MATHS),
    sd=sd(MATHS)
    ) %>%
  mutate(se=sd/sqrt(n-1))
knitr::kable(head(my_sum), format = 'html')
```

The data can then be plotted using the code chunk below.

```{r}
ggplot(my_sum) +
  geom_errorbar(
    aes(x=reorder(RACE, -mean), 
        ymin=mean-1.96*se, 
        ymax=mean+1.96*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  labs(x = "Maths score",
       title = "95% confidence interval of mean maths score by race")
```

An interactive version of this plot can be created.

```{r}
shared_df = SharedData$new(my_sum)

bscols(widths = c(4,8),
       ggplotly((ggplot(shared_df) +
                   geom_errorbar(aes(
                     x=reorder(RACE, -mean),
                     ymin=mean-2.58*se, 
                     ymax=mean+2.58*se), 
                     width=0.2, 
                     colour="black", 
                     alpha=0.9, 
                     size=0.5) +
                   geom_point(aes(
                     x=RACE, 
                     y=mean, 
                     text = paste("Race:", `RACE`, 
                                  "<br>N:", `n`,
                                  "<br>Avg. Scores:", round(mean, digits = 2),
                                  "<br>95% CI:[", 
                                  round((mean-2.58*se), digits = 2), ",",
                                  round((mean+2.58*se), digits = 2),"]")),
                     stat="identity", 
                     color="red", 
                     size = 1.5, 
                     alpha=1) + 
                   xlab("Race") + 
                   ylab("Average Scores") + 
                   theme_minimal() + 
                   theme(axis.text.x = element_text(
                     angle = 45, vjust = 0.5, hjust=1)) +
                   ggtitle("99% Confidence interval of average /<br>maths scores by race")), 
                tooltip = "text"), 
       DT::datatable(shared_df, 
                     rownames = FALSE, 
                     class="compact", 
                     width="100%", 
                     options = list(pageLength = 10,
                                    scrollX=T), 
                     colnames = c("No. of pupils", 
                                  "Avg Scores",
                                  "Std Dev",
                                  "Std Error")) %>%
         formatRound(columns=c('mean', 'sd', 'se'),
                     digits=2))
```

This data can also be visualized using ggdist using the following code.

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_gradientinterval(   
    fill = "skyblue",      
    show.legend = TRUE     
  ) +                        
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Gradient + interval plot")
```

Hypothetical Outcome Plots can also be used.

```{r}
ggplot(data = exam, 
       (aes(x = factor(RACE), 
            y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, 
    width = 0.05), 
    size = 0.4, 
    color = "#0072B2", 
    alpha = 1/2) +
  geom_hpline(data = sampler(25, 
                             group = RACE), 
              height = 0.6, 
              color = "#D55E00") +
  theme_bw() + 
  transition_states(.draw, 1, 3)
```

# Funnel Plot

In this section, COVID-19_DKI_Jakarta will be used. The data was downloaded from [Open Data Covid-19 Provinsi DKI Jakarta portal](https://riwayat-file-covid-19-dki-jakarta-jakartagis.hub.arcgis.com/). The goal of this exercise is to compare the cumulative COVID-19 cases and death by sub-district (i.e. kelurahan) as at 31st July 2021, DKI Jakarta.

The code below imports the data into R.

```{r}
covid19 <- read_csv("data/COVID-19_DKI_Jakarta.csv") %>%
  mutate_if(is.character, as.factor)
```

There are multiple ways to create a funnel plot using the tools at hand. Below are a few of the options available.

::: panel-tabset
### FunnelPlotR

```{r}
funnel_plot(
  numerator = covid19$Death,
  denominator = covid19$Positive,
  group = covid19$`Sub-district`,
  data_type = "PR",   
  xrange = c(0, 6500),  
  yrange = c(0, 0.05),
  label = NA,
  title = "Cumulative COVID-19 Fatality Rate by Cumulative Total Number of COVID-19 Positive Cases", #<<           
  x_label = "Cumulative COVID-19 Positive Cases", #<<
  y_label = "Cumulative Fatality Rate"  #<<
)
```

### ggplot2

```{r}
df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  mutate(rate.se = sqrt((rate*(1-rate)) / (Positive))) %>%
  filter(rate > 0)
fit.mean <- weighted.mean(df$rate, 1/df$rate.se^2)
number.seq <- seq(1, max(df$Positive), 1)
number.ll95 <- fit.mean - 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul95 <- fit.mean + 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ll999 <- fit.mean - 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul999 <- fit.mean + 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
dfCI <- data.frame(number.ll95, number.ul95, number.ll999, 
                   number.ul999, number.seq, fit.mean)
p <- ggplot(df, aes(x = Positive, y = rate)) +
  geom_point(aes(label=`Sub-district`), 
             alpha=0.4) +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll999), 
            size = 0.4, 
            colour = "grey40") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul999), 
            size = 0.4, 
            colour = "grey40") +
  geom_hline(data = dfCI, 
             aes(yintercept = fit.mean), 
             size = 0.4, 
             colour = "grey40") +
  coord_cartesian(ylim=c(0,0.05)) +
  annotate("text", x = 1, y = -0.13, label = "95%", size = 3, colour = "grey40") + 
  annotate("text", x = 4.5, y = -0.18, label = "99%", size = 3, colour = "grey40") + 
  ggtitle("Cumulative Fatality Rate by Cumulative Number of COVID-19 Cases") +
  xlab("Cumulative Number of COVID-19 Cases") + 
  ylab("Cumulative Fatality Rate") +
  theme_light() +
  theme(plot.title = element_text(size=12),
        legend.position = c(0.91,0.85), 
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.background = element_rect(colour = "grey60", linetype = "dotted"),
        legend.key.height = unit(0.3, "cm"))
fp_ggplotly <- ggplotly(p,
  tooltip = c("label", 
              "x", 
              "y"))
fp_ggplotly
```
:::
